---
title: "p8105_hw1_rr3423"
author: "Ragyie Rawal"
date: 2021-09-25
output: github_document
---

```{r setup, include = FALSE}
library(tidyverse)
```


## PROBLEM 1 

```{r df_create}
# Setting seed  
set.seed(1)

# Creating a dataframe
prob_one_df = tibble(
  random_sample = rnorm(10),
  vec_logical = random_sample > 0, 
  vec_character = c("a", "b", "c", "d", "e", "f", "g", "h", "i", "j"),
  vec_factor = factor(c("low", "medium", "high", "low", "medium", "high", "low", "medium", "high", "low" ))
)

knitr::kable(prob_one_df)
```

```{r means}
# taking the mean of each variable in the dataframe
mean_random_sample = mean(pull(prob_one_df, random_sample))

mean_vec_logical = mean(pull(prob_one_df, vec_logical))

mean_vec_character = mean(pull(prob_one_df, vec_character))

mean_vec_factor = mean(pull(prob_one_df, vec_factor))
```

When attempting to take the mean of each variable in the dataframe, it was possible to successfully take the mean of the random sample variable and the mean of the logical variable. It was not possible to take the mean of the character variable or the mean of the factor variable. 

```{r converting_variables, message = FALSE, warning = FALSE}
# Applying the as.numeric function to the logical, character, and factor variables 
convert_logical = as.numeric(pull(prob_one_df, vec_logical))

convert_character = as.numeric(pull(prob_one_df, vec_character))

convert_factor = as.numeric(pull(prob_one_df, vec_factor))
```

Converting the logical variable works because using the as.numeric function converts the "TRUE" and "FALSE" logical values to "1" and "0" numeric values. This helps explain why it is possible to take the mean of the logical variable because the logical values can be converted to numeric values..

Converting the character variable using the as.numeric function does not work because there are only character strings in the character variable, and character strings cannot directly be converted to numeric values. This is why there is a warning message that "NAs introduced by coercion" because R failed to convert the character strings to numeric values. This helps explain why it is not possible to take the mean of the character variable.

Converting the factor variable using the as.numeric function does not give the expected output. This conversion only provides the factor levels, but not the numeric values contained in the factor vector. This is because the factor variable is in character format. This explains why it is not possible to get the mean of the factor variable.


## PROBLEM 2 

```{r penguins_dataset}
# loading dataset 
data("penguins", package = "palmerpenguins")
```

* The **important variables** in the penguins dataset are `r names(penguins)`.

* The **species** in the penguins dataset are `r unique((pull(penguins, species)), na.rm = TRUE)`. 

* The **islands** in the penguins dataset are `r unique((pull(penguins, island)), na.rm = TRUE)`. 

* The **sex** in the penguins dataset are `r unique(na.omit(pull(penguins, sex)))`.

* The **years** in the penguins dataset are `r unique((pull(penguins, year)), na.rm = TRUE)`.

* The **mean** of the bill length is `r round(mean(pull(penguins, bill_length_mm), na.rm = TRUE), digits = 2)`. The **standard deviation** of the bill length is `r round(sd(pull(penguins, bill_length_mm), na.rm = TRUE), digits = 2)`. 

* The **mean** of the bill depth is `r round(mean(pull(penguins, bill_depth_mm), na.rm = TRUE), digits = 2)`. The **standard deviation** of the bill depth is `r round(sd(pull(penguins, bill_depth_mm), na.rm = TRUE), digits = 2)`. 

* The **mean** of the flipper length is `r round(mean(pull(penguins, flipper_length_mm), na.rm = TRUE), digits = 2)`. The **standard deviation** of the flipper length is `r round(sd(pull(penguins, flipper_length_mm), na.rm = TRUE), digits = 2)`. 

* The **mean** of the body mass is `r round(mean(pull(penguins, body_mass_g), na.rm = TRUE), digits = 2)`. The **standard deviation** of the body mass is `r round(sd(pull(penguins, body_mass_g), na.rm = TRUE), digits = 2)`.

* The **number of rows** in the penguins dataset is `r nrow(penguins)`.

* The **number of columns** in the penguins dataset is `r ncol(penguins)`. 

```{r scatterplot_penguins, message = FALSE, warning = FALSE}
# making scatterplot for flipper length and bill length
ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) + geom_point() + ggtitle("Flipper Length vs. Bill Length")

# exporting scatterplot to project directory 
ggsave("p2_scatter_plot.pdf", height = 4, width = 6)
```
